---
title: "DATA1901 International Airlines Operating in Australia"
author: "Scarlett Foulsham"
date: "University of Sydney | DATA1901 | November 2021"
output:
  html_document:
    code_download: true
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true 
    theme: sandstone
    highlight: tango
---

```{css newstyles, echo=FALSE}
h1, .h1, h2, .h2, h3, .h3 { /* Add space before headings: */
    margin-top: 56px;
}
h1, h2 { /* add border to h1 and h2 */
  border-bottom: solid 1px #666;
}
h2 { /* Resize header 2: */
  font-size: 22px;
}
h3 { /* Resize header 2: */
  font-size: 18px;
}
body { /* Make main text colour black */
  color: black;
}
a { /* Link colours */
  color: blue;
}
.tocify { /* Some toc fixes*/
  width: 100% !important;
  border: none; /* remove border */
}
.tocify-header { /* fix for horrible indent in toc */
  text-indent: initial;
}

```

# Recommendation
Client: fellow airline industry workers

Based on findings within this report, we advise increased travel advertisement during the 
months leading up to September and December, an inquiry into the Qantas Airways buisness 
model, maintenance checks for route SYD-AKL, and research into what makes Australian-New 
Zealand relations so successful.

# Initial Data Analysis
Let's load in the data and some packages.
```{r, message=FALSE}
data = read.csv("flights.csv")

library(tidyverse)
library(plotly)
library(gapminder)
library(wordcloud)
library("RColorBrewer")
library(plotly)
library(hrbrthemes)
library(data.table)
library(reactable)
```

Let's check the structure and ensure no `NA` values.
```{r}
str(data)
sum(is.na(data))
```

Let's move onto exploring the data. What are the most popular months for travel within Australia? 
We want to increase advertising leading up to these months to maximize potential revenue.
```{r, echo=FALSE, results="hide"}
# What months are most popular for travel to Australia?

# Finds the total number of flights for each month
month_flight_counts = aggregate(x = data$All_Flights,
                                by = list(data$Month_num),
                                FUN = sum) 

# Finds the percentage of total flights that occur in each month
# Will use this information when making my bar plot
records_per_month <- list(200495, 173891, 210082, 166419, 164169, 184968,
                          176751, 172710, 200931, 173712, 167062, 220202)
for (x in records_per_month) {
  print((x/2211392)*100)}
```

```{r, echo=FALSE}
# Making the bar plot
# Setting up the order in which I want to x-axis labels to appear
xform <- list(categoryorder = "array",
              categoryarray = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                                "Aug", "Sept", "Oct", "Nov", "Dec"))

# Now making the bar plot
fig <- plot_ly(
  x = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sept", "Oct", "Nov", "Dec"),
  y = c(9.066, 7.863, 9.4910, 7.526, 7.424, 8.364, 7.993, 7.810, 9.086, 7.855, 7.555, 9.958),
  text = c("9.066% of total flights", "7.863% of total flights", "9.4910% of total flights",
          "7.526% of total flights", "7.424% of total flights", "8.364% of total flights",
          "7.993% of total flights", "7.810% of total flights", "9.086% of total flights",
          "7.855% of total flights", "7.555% of total flights", "9.958% of total flights"),
  type = "bar",
  marker = list(color = c("lightblue", "lightblue", "#ffa07a", "lightblue", "lightblue", "lightblue",
                          "lightblue", "lightblue", "#ffa07a", "lightblue", "lightblue", "#ffa07a"))) %>% layout(xaxis = xform)

# Adding titles and printing the final bar plot
fig <- fig %>% layout(title = "Most Popular Months for Travel to Australia",
                      xaxis = list(title = "Months"), yaxis = list(title = "Percentage of Total Flights"))
fig
```
See the three most popular months highlighted in orange: `December` followed by `March` then `September`.


What airlines are most successful? We want to study their business models and understand 
why flyers choose certain airlines over others operating in their home country. See word 
cloud below for an overall visual understanding and the table of the top 10 most successful 
airlines for a more concrete understanding.
```{r, echo=FALSE}
# What airlines are most successful?

# This finds the total number of flights for each airline
# Again will use this information to make my word cloud
airline_flight_counts = aggregate(x = data$All_Flights,
                                  by = list(data$Airline),
                                  FUN = sum) 

# Renaming the column names 
names(airline_flight_counts)[1] <- "AirlineName"
names(airline_flight_counts)[2] <- "Total_Number_of_Flights"
```

```{r, warning=FALSE, echo=FALSE}
# Now let's create the word cloud
# Added some colors and different specifics  
wordcloud(words=airline_flight_counts$AirlineName, freq=airline_flight_counts$Total_Number_of_Flights,
          colors=brewer.pal(10, "Paired"), scale=c(6, 1), 
          max.words=35, rot.per=0.3)
```

```{r, echo=FALSE}
# Making a table so readers can read what the word cloud visualises 
sort_data <- airline_flight_counts[order(airline_flight_counts$Total_Number_of_Flights),] 
tail_data <- tail(sort_data, 10)
tail_dataframe <- as.data.frame(tail_data)
reactable(tail_dataframe)
```
We can see that some of the highest performing airlines are `Qantas Airways`, `Jetstar`, `Virgin Australia`, 
`Air New Zealand`, `Emirates`, `Cathay Pacific Airways`, and `Garuda Indonesia`.


What are the top five most frequent routes? We want to watch maintenance checks for planes 
flying these routes in particular and ensure availability year round for these routes.
```{r, echo=FALSE, results="hide"}
# What's the most popular route?

# Quickly finding the total number of flights for each route 
route_flight_counts = aggregate(x = data$All_Flights,
                                by = list(data$Route),
                                FUN = sum) 

# Printing only those greater than 20,000
for (i in c(route_flight_counts$x)){
  if(i>20000){
    print(i)
  }}
```

```{r, echo=FALSE}
# Now let's make the lollipop plot
# Setting up the x and y values 
# Creating a dataframe 
x <- c("AKL-BNE", "AKL-MEL", "AKL-SYD", "BNE-AKL", "CHC-SYD", "DPS-PER",
           "HKG-SYD", "MEL-AKL", "MEL-SIN", "PER-DPS", "PER-SIN", "SIN-MEL",
           "SIN-PER", "SIN-SYD", "SYD-AKL", "SYD-CHC", "SYD-HKG", "SYD-SIN")
y <- c(25128, 28063, 50657, 25125, 20093, 27681,
            24629, 27867, 20453, 27614, 31620, 20472,
            31628, 23748, 50801, 20091, 24619, 23755)
routes.data <- data.frame(x, y)

# Ordering them based on y
routes.data <- routes.data %>% 
  arrange(y) %>% 
  mutate(x=factor(x,x))

# Let's now create the actual plot
ggplot(routes.data, aes(x=x, y=y)) +
  geom_segment(aes(x=x, xend=x, y=0, yend=y),
    color="skyblue", 
    size=ifelse(routes.data$x %in% c("SYD-AKL", "AKL-SYD", "SIN-PER", "PER-SIN", "AKL-MEL"), 1, 0.7)) +
  
  geom_point( color="darkorange", 
              size=ifelse(routes.data$x %in% c("SYD-AKL", "AKL-SYD", "SIN-PER", "PER-SIN", "AKL-MEL"), 4, 2), 
              alpha=0.6) +
  xlab("Travel Routes") + ylab("Number of Flights") +
  ggtitle("Most Popular Travel Routes for International Airlines") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank())
```

Take a look at the top five emphasized lines. The routes are from Sydney, Melbourne, and Perth all to `Auckland, New Zealand`.


What international countries visit Australia most frequently? We want to create a personalised 
experience for our most frequent customer base and understand how to attract flyers from other countries.
```{r, echo=FALSE, eval=FALSE}
# What international countries visit Australia the most?

# This finds the number of flights for each country
country_flight_counts = aggregate(x = data$All_Flights,
                                  by = list(data$Port_Country),
                                  FUN = sum)  

# Ordering the previous aggregate and looking at only the greatest values which are at the end
barplot_data <- country_flight_counts[order(country_flight_counts$x),] 
tail_data <- tail(barplot_data, 15)
reactable(tail_data)
```

```{r, echo=FALSE}
# Let's create the bar plot now 
# Establishing the order I want the x values to appear
xform <- list(categoryorder = "array",
              categoryarray = c("Philippines", "Vietnam", "Papua New Guinea", "UK", "Fiji", "Japan",
                                "Thailand", "China", "Malaysia", "United Arab Emirates", "Hong Kong (SAR)",
                                "USA", "Indonesia", "Singapore", "New Zealand"))

# Creating the plot
fig <- plot_ly(
  x = c("Philippines", "Vietnam", "Papua New Guinea", "UK", "Fiji", "Japan",
        "Thailand", "China", "Malaysia", "United Arab Emirates", "Hong Kong (SAR)",
        "USA", "Indonesia", "Singapore", "New Zealand"),
  y = c(26804, 27736, 48946, 50413, 59038, 78667, 92788, 107252,
        112948, 127519, 143359, 152913, 159915, 291164, 534734),
  type = "bar",
  marker = list(color = c("mistyrose", "mistyrose", "mistyrose", "mistyrose",
                          "mistyrose", "mistyrose", "mistyrose", "mistyrose", 
                          "mistyrose", "mistyrose", "mistyrose", "mistyrose", 
                          "salmon", "salmon", "salmon"))) %>% layout(xaxis = xform)

# Adding some basics like titles and printing the plot
fig <- fig %>% layout(title = "Most Frequent International Countries Visiting Australia",
         xaxis = list(title = "Countries"),
         yaxis = list(title = "Number of Flights"))
fig
```

We can see that `New Zealanders` are the vast majority of our international visitors, followed by `Singapore` and `Indonesia`.
Let's move onto some more formula research questions.

# Research Questions 

## Is there a relationship between most seats available and most frequent route? 
We want to ensure that airlines are offering high seat availability for the 
most frequent routes to maximize profit. Let's explore through an independence test.

H0: Most frequent route is independent of seat count.

H1: Most frequent route is *not* independent of seat count.

Assumptions: For expected counts, none are empty and no more than 20% have counts of <5.
```{r, warning=FALSE}
# Independence test
mytable <- table(data$Route, data$Max_Seats)
results <- chisq.test(mytable)
print(results)
```

We observe a p-value of `2.2e-16` which is extremely far below 0.05. Therefore our 
*statistical conclusion* is that we reject H0 as p < 0.05 and our *scientific conclusion* 
is that we can say with almost no doubt that most frequent route is not independent of
seat number meaning they have some relationship. See relationship visualised below.

```{r, echo=FALSE, warning=FALSE}
# Let's create a scatterplot to visualise the relationship that we just confirmed is present

# We have previously established a table for routes by flights
names(route_flight_counts)[1] <- "Route"
names(route_flight_counts)[2] <- "Number_of_Flights"

# Let's now make one for seats by route
route_seats_counts = aggregate(x = data$Max_Seats,
                               by = list(data$Route),
                               FUN = sum) 
names(route_seats_counts)[1] <- "Route"
names(route_seats_counts)[2] <- "Number_of_Seats"

# Let's make the table into a dataframe
# One column for seats and another for frequency of flights (all based on routes)
jointdataset <- merge(route_seats_counts, route_flight_counts, by="Route")

# Creating the scatterplot
ggplot(jointdataset, aes(x=Number_of_Seats, y=Number_of_Flights)) + 
  geom_point(size=2, color="lightblue") +
  ylab("Number of Flights per Route") +
  xlab("Number of Seats") + 
  ggtitle("How Seat Number Increases with Route Frequency") +
  theme_light()
```

As the frequency of a `route being flown increases`, it appears that the `number of seats offered also increases`.

## Are flights increasing linearly with year? 
If not, this could be an indication that we need to improve service and increase marketing. Let's explore through a linear model.

```{r, echo=FALSE}
# Finding the total number of flights for each year 
flights_by_year = aggregate(x = data$All_Flights,
                            by = list(data$Year),
                            FUN = sum) 

# Renaming the columns
names(flights_by_year)[1] <- "Year"
names(flights_by_year)[2] <- "Number_of_Flights"

# Printing the first couple rows 
head(flights_by_year)
```

```{r}
# Creating the linear model 
fit <- lm(Number_of_Flights ~ Year, data=flights_by_year)
summary(fit)

# Finding the correlation coefficient 
cor(flights_by_year$Year, flights_by_year$Number_of_Flights)
```

```{r, echo=FALSE}
# Making a residual plot
ggplot(fit, aes(x=.fitted, y=.resid)) +
  geom_point(color="darkorange") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "sienna1") +
  ggtitle("Residual vs Fitted") +
  xlab("Predicted Values") + 
  ylab("Standarised Residuals") +
  theme_light()
```

Even though the linear model has a moderately strong correlation coefficient of nearly `0.6`,
the residual plot demonstrates a clear trend so we deem the model to be unreliable. 
Let's switch to exploring through a line plot.

```{r, echo=FALSE}
# Now making a linear plot instead
x <- flights_by_year$Year
y <- flights_by_year$Number_of_Flights
data <- data.frame(x, y)
fig <- plot_ly(data, x = ~x, y = ~y, type = 'scatter', 
               mode = 'lines', color=I("lightseagreen"))
fig <- fig %>% layout(title = "Number of Flights by Year",
         xaxis = list(title = "Number of Flights"),
         yaxis = list(title = "Year"))
fig
```

We can see an `increasing trend` until late 2017 when COVID dramatically reduced 
the number of flights taken. This indicates the relative success of the airlines 
industry and raises the question of how we should encourage international travel 
to Australia post COVID.

For further explanation, please see our explanatory video!
